\documentclass[../FinalThesis.tex]{subfiles}
\begin{document}

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
\thispagestyle{empty}
\vspace{-1cm}
\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil

\newpage

This work would not have been possible without help and support from countless
people and organizations. I'd like to start by thanking the Ministry of
Environment and Tourism of Botswana for generously granting permission to
conduct research for this project. Financial support for this thesis was
provided by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust,
Forschungskredit of the University of Zürich (no. FK-21-091), a Swiss National
Science Foundation grant (no. 31003A\_182286) awarded to Arpat Ozgul, and a
Swiss National Science Foundation grant (no. 310030\_204478) awarded to Gabriele
Cozzi.

I am particularly grateful to my supervisor, Gabriele Cozzi, who has initiated
and managed the African wild dog dispersal project over the past decade and has
been an outstanding mentor. His leadership and dedication to research and
conservation were absolutely instrumental to this work. He provided unstinting
support for my doctoral thesis and has significantly contributed to the success
of each chapter. Gabriele played a pivotal role in securing both material and
salary funds for my research, and eventually included me under his own research
grant. Moreover, he introduced me to the challenges and rewards of fieldwork in
one of this planet's most pristine environments. Through his training, I
acquired a diverse set of skills, from tracking and monitoring large carnivores,
navigating and driving in the African Savannah, sedating and equipping animals
with GPS satellite collars, to fixing Land Rovers. Gabriele's constant
supervision ensured my scientific progress, and he continuously provided timely
guidance whenever I veered off course. His unwavering motivation and belief in
my abilities kept me energized, especially during moments of self-doubt. Even
beyond academic support, Gabriele spared no expense and happily assisted Manuela
and me in our endeavor of finding a Land Rover for our upcoming adventures in
Africa. Besides our working relationship, Gabriele and I have developed an integral
friendship, which hopefully continues to grow in the future. Thanks for the
countless BBQs, beers, and Italian fine dines. I don't think anybody can wish
for a better PhD supervisor.

I extend my gratitude to Arpat Ozgul, whose belief in my scientific abilities
and provision of funds enabled me to embark on my PhD as part of his research
group. Arpat was an excellent supervisor, and his scientific rigor significantly
advanced my understanding of ecology and propelled my academic growth. It was a
privilege to be part of his research group and I found tremendous pleasure in
participating in the various group-retreats, potlucks, and other social
gatherings.

I am immensely grateful to Dominik Behr, who has accompanied my academic career
from the very beginning. Dominik demonstrated to me that transitioning into the
field of ecology was not only feasible, but also highlighted the personal
fulfillment this pursuit may bring about. Under his mentorship, I conducted my
master's thesis on dispersing AWDs, which ultimately laid the foundation for
this PhD. He also infected me with the African virus, giving me an unhealthy
passion for African wild dogs and African wildlife in general. Beyond academic
mentorship, Dominik has continuously inspired me to embrace creativity when
crafting PowerPoint presentations, scientific figures, or when taking photos of
wildlife. It's been an incredible journey, and I've enjoyed each and every of
our many fruitful discussions.

I am deeply thankful to have had an opportunity to collaborate with John
Fieberg, whose expertise was invaluable to my academic journey. John is among
the most humble yet inspiring people I have ever met. His ability to return any
email with an in-depth response within less than 24 hours is one in a kind and
makes collaborating with him an absolute blast. If one imagines a statistics
professor as being an uncoordinated nerd, beware of John's ability to throw a
Frisbee at the speed of light; it will prove you wrong. It was a pleasure
organizing and teaching a movement ecology workshop with John in Faido, and I am
looking forward to organizing another one in the future.

I extend my sincere appreciation to thank all my committee members, namely Lukas
Keller, Arpat Ozgul, Juan Manuel Morales, John ``Tico'' McNutt, and Gabriele
Cozzi for their valuable insights and stimulating discussions during our
committee meetings. Their collective expertise and guidance tremendously
facilitated a series of focused and coherent research chapters. I'd particularly
like to thank Lukas for adopting me in his research group without hesitation.

I would like to express special thanks to Florian Hodel for his contagious
ingenuity and infectious humor, which never failed to brighten even the dullest
moments. Florian's creativity has sparked an incredible amount of interesting
ideas and has his friendship has been invaluable to me. It was a shame we only
overlapped for such a short time at the University, although I have a notion
that it was for the better of my productivity. Maybe $\pi$-ranha will provide
further opportunities for collaborations in the future. I also want to thank his
wife, Rahel, who was a source of strength and probably provided more life-advice
than she is aware of. The joy shared with you both has been unparalleled; I even
suffer from soar jaw-muscles from all the laughter whenever we spend time
together.

I am also grateful to all current and former Popecol members, including
Benedetta, Che, Daniela, Dechen, Dilsad, Esin, Eva, Lorena, Lukas, Lorena,
Louis, Monica, Omar, Patrick, Philip, Rasmitah, Tis, and Vera-Lu for an
unforgettable time and many exciting moments. I'd like to especially thank two
master students that I co-supervised, Megan Robinson and Rapha{\"e}l Destriau,
for their effort and motivation which have laid the foundation for future works.
It's rare to supervise students with such drive and dedication, and it's been a
pleasure being part of their exciting works.

Further gratitude is due to the University's secretaries and technical support,
who ensured that all organizational tasks were taken care off; something I'm
terribly bad at. Many thanks to Jacqui, Judith, Ursina, Debra, Susanne, Isabel,
Michel, and Tina.

Thanks also to Johannes Signer and Theo Michelot for sharing their insights on
how to simulate movement from fitted step-selection functions, and to Guillaume
Bastille-Rousseau for helping me with computing network metrics from GPS data.

Furthermore, I want to extend my gratitude to Glauco Camenisch, who assisted at
countless wild dog immobilizations and has always provided an ear when I needed
one. Experiencing the bush together with you and Gabs was a unique privilege and
among the most fun I've had in Africa thus far.

In Botswana, a heartfelt ``Thank You'' goes to Tico and Lesley for establishing
Dog Camp and for the passion and love that they keep pouring into this amazing
research project. Through Wild Entrust, you have created something truly
extraordinary that holds the promise of providing a future where humans and
wildlife can coexist. It is an honor to be part of this endeavor, and one can
only aspire to contribute to its success in the future.

Special thanks to Peter Brack for showing me how to cross water with a Land
Rover (a proper lake, in fact), how to hot-wire a Land Rover, for organizing a
walking safari (let's do another one!), and for the countless inspiring
discussions we had in camp. Similarly, I'd like to thank his wife, Nicola Hawes,
for accommodating me on various occasions in their wonderful ``termite mound''
house in town. Further thanks go to Megan Claase, for exemplifying true
dedication and passion for conservation and for serving as a fieldwork role
model. Our expedition to David's Kingdom marks one of my favorite trips in
Africa to date. Further thanks go to Whitney Hansen, for bringing me on some
incredible wild dog follows and for introducing me to the imperial system
(``come within a whale of me and say that again''). Thanks to Emily Bennitt,
Mark and Caroll Swaffield for providing me with shelter and food in Maun, to Rob
Jackson for providing veterinary services, and to Mary Hastag from Abercrombie
\& Kent for organizing several trips into the Okavango Delta and for initiating
an exciting collaboration that hopefully continues into the future. Gratitude is
also due to all research assistants in Botswana, including Rio Dipuo, John
Neelo, Megan Robinson, Alex Dibnah, and Mara Zali, for relentlessly collecting
data in the field, as well as to all staff members in camp, including Allen, for
providing mechanical support and fixing our Land-Rovers and for his musical
irradiation in the shower, to Fine and Tea for preparing the most elaborate
meals with the least elaborate items, to Borden, Opi, and Obie for keeping camp
clean and tidy, and to Last, for tirelessly fixing our continuously punctured
tires.

Special thanks go to our former research assistant Aobakwe Sengate (a.k.a. AB),
for being an incredible field-mentor and for filling my knowledge gaps with
regard to the African wildlife. Though our time in the field was short, every
trip with AB was exciting, even if it only was to check cameratraps. AB's
abilities to read the environment and track animals are second-to-none, and his
never-ending curiosity and willingness to investigate even the faintest evidence
of large predators were inspiring and informative. He taught me some of the most
useful skills, and I've enjoyed some of the most memorable sightings with him.
I'll never forget the look on his face when he realized a lioness stole a
leopard's kill just behind his tent in camp.

My deepest gratitude goes to my family, including my parents, Rahel and Denis,
and my grandparents Elizabeth, Hans-Peter, and Josephine. They have always been
there for me and provided me with all the psychological and financial stability
that I needed throughout this journey. Similarly, I'd like to thank my
girlfriend's family, namely her parents Yvonne and Max, as well as her
grandparents Willy and Annula. I also want to thank her grandfather Ernst, who
sadly passed away recently. Being embraced by such a wonderful social
environment is a gift beyond measure, and I am incredibly grateful for the love
and support I have received from both families.

Finally, I want to thank my girlfriend Manuela, who supported me through all of
my undertakings, no matter how devious some of them must have sounded to her.
Despite her own demanding schedule, Manuela consistently found the time and
energy to uplift and motivate me when I needed it most. Manuela was not only at
peace with my passion for Africa, but got infected with an unhealthy enthusiasm
herself. Please excuse that I have spent several months in the field, leaving
you and our cat Nala behind, and thank you for crunching through the countless
cameratrap images (there will be more...). It is not a given to find a partner
that can cope with a scientist's lifestyle, and I appreciate your unconditional
support. I can't wait to explore the African wilderness further with you in our
Landy!

\end{document}
