\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Dispersal Through Seasonal Landscapes}

% \title{Step by Step: The Utility of Dispersal Simulations to Assess Landscape
% Connectivity}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,\S} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Gabriele Cozzi\textsuperscript{1,2} \and
  Dominik M. Behr\textsuperscript{1,2}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
\date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation, Private Bag 13, Maun,
Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\vspace{4cm}

\textbf{Running Title:} Release the Dogs! Simulating Wild Dog Dispersal to
Assess Landscape Connectivity

\vspace{0.5cm}

\textbf{Keywords:} dispersal, simulation, movement, integrated step selection,
Kavango-Zambezi Transfrontier Conservation Area, landscape connectivity, Lycaon
pictus

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
This is the abstract.
\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

\onehalfspacing
\tableofcontents
\doublespacing

% Change page numbering
\newpage
\pagenumbering{arabic}

% Create linenumbers
\linenumbers

\section{Introduction}

% Taken from Zeller.2020
Connectivity for wildlife is typically modeled in static environments and
estimated for a single snapshot in time (e.g., [6,7]). However, wildlife
responses to landscapes are inherently dynamic, as are the landscapes themselves
(e.g., [8])

Modeling connectivity in a static environment may mask the importance of connectivity for
maintaining wildlife population dynamics and may result in identifying corridors that are only
periodically functional [15,16,18]

\section{Methods}

\section{Results}

\section{Discussion}

\section{Authors' Contributions}
D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
first draft of the manuscript and all authors contributed to the drafts at
several stages and gave final approval for publication.

\section{Data Availability}
GPS movement data of dispersing coalitions will be made available on dryad at
the time of publication. Access to all R-scripts for our analyses is provided
through Github.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
to Johannes Signer for assisting with the simulation algorithm. This study was
funded by Basler Stiftung für Biologische Forschung, Claraz Foundation, Idea
Wild, Jacot Foundation, National Geographic Society, Parrotia Stiftung, Stiftung
Temperatio, Wilderness Wildlife Trust Foundation, Forschungkredit der
Universität Zürich, and a Swiss National Science Foundation Grant
(31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}
