---
title : 'Example 2: Dynamic + Model Approach with Real Data'
author: "David D. Hofmann, Gabriele Cozzi, John Fieberg"
date  : "2024-01-30"
output:
  html_document:
    toc      : true
    toc_depth: 2
---
```{r setup, echo = F}

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

## Purpose
This example analysis serves to illustrate the application of the _dynamic +
model_ approach with real data collected on a spotted hyena (*Crocuta crocuta*)
called "Apollo". First, we'll do some slight pre-processing of the GPS data.
Since missingness is really low in this dataset, we'll artificially increase it
to a higher level. We will then apply the _dynamic + model_ approach to estimate
habitat and movement parameters.

## Disclaimer
In this script, we'll utilize a set of custom functions that are stored in the
`Functions.R` file. They are used to compute step-metrics, generate
random-steps, and fit step distributions dynamically. Most of the functions can
be replaced by much more generalized functions from the `amt` R-package.
However, `amt`'s `random_steps()` function doesn't cope well with sampling
random steps from distributions fit to different step-durations. We'll therefore
use of our custom code. If you'd like to apply it to your own data, inspect the
functions closels and verify that they suit your purposes.

## Preamble
Let's start by preparing our R-Session and loading all necessary packages. We
also load the custom functions stored in the `Functions.R` file.

```{r, warning = F, message = F}

# Set notation standards
options(scipen = 999)

# Clear R's brain
rm(list = ls())

# Load required packages
library(tidyverse)     # For data wrangling
library(sf)            # For plotting spatial features
library(lubridate)     # To handle dates and times
library(raster)        # For spatial data manipulation
library(amt)           # To fit step-length and turning-angle distributions
library(survival)      # To fit conditional logistic regression model
library(ggh4x)         # For nested ggplot facets

# Load custom functions
source("02_R-Scripts/Functions.R")

```

## Pre-Processing
We start by loading the GPS data of a spotted hyena (*Crocuta crocuta*) and some
covariate layers.

```{r}

# Load hyena gps data and covariates
dat <- read_rds("03_Data/Hyena_GPS.rds") %>% subset(DOP < 10)
cov <- stack("03_Data/Hyena_Covariates.tif")

```

The hyena GPS data should have been collected every 2 hours, starting from 16:00
and until 04:00 o'clock each day. There usually is also a fix collected at 10:00
o'clock. Since hyenas are mainly nocturnal, we're only going to focus on the
nightly data and drop the 10:00 o'clock fix. In very rare cases, two fixes fall
on the same hour. We'll drop such "duplicates" to avoid having step-durations
that are less than 2 hours.

```{r}

# Keep only fixes that were collected on the desired hour and drop potential
# duplicates
dat <- dat %>%
  mutate(timestamp = round_date(timestamp, "hour")) %>%
  subset(hour(timestamp) %in% c(16, 18, 20, 22, 0, 2, 4)) %>%
  distinct(id, timestamp, .keep_all = T)

# Assess the regular step-duration of our data
samp <- dat %>%
  computeDurations() %>%
  count(duration)
samp

```

If you take a look at the durations between subsequent fixes, you'll find that
most fixes are taken at 2-hourly intervals, with only few 4- or 6-hourly
periods. This means that the data is pretty complete (i.e. there is relatively
low missingness). To showcase the applicability of the *dynamic+model* approach
we'll introduce a slightly higher level of missingness by removing a fraction of
the collected GPS fixes.

```{r}

# Incrase missingness a bit
dat_rarified <- dat %>%
  rarifyData(missingness = 0.15) %>%
  mutate(
      step_id     = 1:n()
    , step_number = 1:n()
  )

# Compare number of fixes
cbind(Original = nrow(dat), Rarified = nrow(dat_rarified))

```

Let's visualize everything, including the spatial covariate layers.

```{r}

# Visualize covariates and the rarified track (normalization is only for
# visualization purposes)
cov_df <- cov %>%
  as.data.frame(xy = T) %>%
  pivot_longer(Water:Trees, names_to = "Covariate", values_to = "Value") %>%
  group_by(Covariate) %>%
  mutate(Value = (Value - min(Value, na.rm = T)) / (max(Value, na.rm = T) - min(Value, na.rm = T)))
ggplot() +
  geom_raster(data = cov_df, aes(x = x, y = y, fill = Value)) +
  geom_path(data = dat_rarified, aes(x = x, y = y, col = timestamp), linewidth = 0.1) +
  scale_fill_viridis_c(breaks = c(0, 0.5, 1), labels = c("Low", "Medium", "High")) +
  scale_color_datetime(low = "red", high = "green", name = "Timestamp") +
  facet_wrap(~ Covariate) +
  coord_equal() +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "gray95", color = "white"))

```

## Fitting Tentative Distributions
Because we rarified the dataset, the time between consecutive datapoints is not
always equal. Let's ensure the regular step-duration is still 2 hours.

```{r}

# Assess the regular step-duration of our data
samp <- dat_rarified %>%
  computeDurations() %>%
  count(duration)
samp

```

As expected, the regular step-duration is 2 hours. However, we also observe that
in a few cases the duration between two fixes can be as long as `r
max(samp$duration)` hours. In regular iSSF, we would only consider bursts within
which step-durations are somewhat comparable. Here, however, we are tolerant
towards unequal step-durations, as we'll later account for differing
step-durations in the model. Consequently, we would like to accommodate for
differing durations by fitting tentative distributions for different
step-durations. Here, allowing for steps of up to 6 hours appears sensible. Let
us fit the associated step-length and turning angle distributions using the
`fitDists()` function.

```{r}

# Fit step-length and turning angle distributions to different durations
dists <- fitDists(dat_rarified
  , durations = c(2, 4, 6) # Return fitted distributions for durations 1-3
  , regular   = 2          # The regular step-duration is 2
  , dynamic   = T          # Fit distributions dynamically
  , resample  = F          # Don't resample the track
  , rarify    = F          # Don't increase missingness
)

# Show the fitted values
dists

```

You can see that we now have different gamma and von Mises parameters for steps
from different durations. We can later use the so fitted distributions to
generate random steps matching different step-durations.

## Generating Random Steps
Before generating random steps, we need to split our data into bursts within
which the step-duration does not exceed 6.

```{r}

# Compute bursts within which the step duration does not exceed our forgiveness
dat_rarified_bursted <- computeBursts(dat_rarified, max_duration = 6)
head(dat_rarified_bursted)

```

In addition, we need to compute step-metrics (step-lengths and turning-angles)
per burst.

```{r}

# Compute step metrics within bursts
dat_rarified_metrics <- computeMetrics(dat_rarified_bursted)
head(dat_rarified_metrics)

```

Finally, we can use the `computeSSF()` function to generate 25 random steps for
each valid step (a valid step has a step-length and a turning-angle associated).
Note that we cna use the dynamically fitted distributions `dists` to match
sampled random steps with the step-durations of observed steps by picking the
`dynamic+model` approach. The function `computeSSF()` will also automatically
drop any invalid steps or steps with a duration that occurs less than 5 times in
the dataset. This is to prevent issues when including the step-duration in the
iSSF model.

```{r}

# Generate random steps
dat_rarified_ssf <- computeSSF(dat_rarified_metrics
  , n_rsteps = 25
  , dists    = dists
  , approach = "dynamic+model"
)
head(dat_rarified_ssf)

```

Let's visualize some of the observed and random steps to ensure everything is
working as expected.

```{r}

# Verify that steps of different durations follow different sl- and
# ta-distributions
dat_rarified_ssf %>%
  dplyr::select(case, duration, sl, relta) %>%
  pivot_longer(sl:relta, names_to = "Metric", values_to = "Value") %>%
  mutate(Type = ifelse(case == 1, "Observed", "Random")) %>%
  ggplot(aes(x = Value, col = as.factor(duration))) +
    geom_density() +
    facet_nested_wrap(~ Type + Metric, scale = "free") +
    scale_color_viridis_d(name = "Step-Duration") +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "gray95", color = "white"))

# Visualize some observed and random steps
dat_rarified_ssf %>%
  mutate(Type = ifelse(case == 1, "Observed", "Random")) %>%
  subset(step_number %in% 1:1000) %>%
  ggplot(aes(x = x, y = y, xend = x_to, yend = y_to, col = as.factor(duration))) +
    geom_segment(linewidth = 0.1) +
    scale_color_viridis_d(begin = 0.3, name = "Duration") +
    theme_minimal() +
    coord_equal() +
    facet_wrap(~ Type) +
    theme(strip.background = element_rect(fill = "gray95", color = "white"))

```

To be able to analyse our trajectory in an iSSF-framework, we need to generate
random steps. We'll use the `amt` package to generate these. However, unlike in
traditional iSSFs, we will generate random steps using tentative distributions
that are matched with the step-durations of observed steps. Let's start by
converting our simulated dataframe into a `track` object (which is actually just
a `data.frame` that contains all elements that `amt` requires to compute
step-metrics).

## Computing Covariates
We still need to compute all covariates that we want include our iSSF model.
This includes spatial covariates at the end-point of each step, as well as
step-metrics. We can use the function `computeCovars` for this.

```{r}

# Compute covariates for all steps (the function will extract spatial covariates
# at the end of each step and also compute log_sl and cos_ta)
dat_rarified_covars <- computeCovars(dat_rarified_ssf
  , covars    = cov
  , extract   = "end"
)
head(dat_rarified_covars)

```

## Running the Model

We are now ready to estimate parameters using conditional logistic regression.
This can be achieved using the `clogit()` function from the `survival` package.
Alternatively, one can use the wrapper `fit_clogit()` from the `amt` package.
Importantly, we will include interactions among the step-descriptors $sl$,
$log(sl)$, and $cos(ta)$ and the step-duration $duration$.

```{r}

# Ensure the step-duration enters as factor
dat_rarified_covars$duration <- as.factor(dat_rarified_covars$duration)

# Fit the model
mod <- clogit(data = dat_rarified_covars, formula = case ~
  + Water                     # Habitat-selection
  + DistanceToWater           # Habitat-selection
  + Trees                     # Habitat-selection
  + Water:duration            # Habitat x Step-Duration
  + DistanceToWater:duration  # Habitat x Step-Duration
  + Trees:duration            # Habitat x Step-Duration
  + sl                        # Movement Kernel
  + log_sl                    # Movement Kernel
  + cos_ta                    # Movement Kernel
  + sl:duration               # Movement x Step-Duration
  + log_sl:duration           # Movement x Step-Duration
  + cos_ta:duration           # Movement x Step-Duration
  + strata(step_id)
)

# Compute confidence intervals
ci <- confint(mod, level = 0.95)

# Put relevant results together
coefs <- summary(mod)$coefficients
coefs <- data.frame(
    Coefficient = rownames(coefs)
  , Estimate    = coefs[, "coef"]
  , SE          = coefs[, "se(coef)"]
  , Z           = coefs[, "z"]
  , LCI         = ci[, 1]
  , UCI         = ci[, 2]
)

# Show results
coefs

```

```{r}
sessionInfo()
```
